SunPy - Python for Solar Physics

### 2014-01-30 1-Introduction.tex

Science is driven by the analysis of data. Modern advances in sensor
technology, combined with the availability of inexpensive storage has led to
rapid increases in amount of data available to scientists in almost
every discipline.  Solar physics is no exception to this trend. For example, NASA's
AIA instrument on the Solar Dynamics Observatory (SDO) satellite, launched in February 2010,
produces over 1 TB of catalogued and stored data per day \citep{aia}. Managing and
analysing this mountain of data
requires increasingly sophisticated software tools.  Key qualities
that these tools should have include robustness, ease of use, a
transparent development history, ease of modification, and conformance
to standards.  Software with these qualities provide a strong
foundation and can support the needs of the community as data
volumes grow and science questions evolve.
%schriste - the opening paragraph still needs work
%ji - I had a go.
%schriste - looks good, made a small fix

SunPy aims to provide a free, open-source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and the scientific Python packages. Python is a general-purpose, 
powerful and easy-to-learn high-level programming language.
% RJH: Can we cut the next sentence?  It seems superfluous and top 10 is rather
% meaningless.
% schriste - i think it's important. it contrasts nicely with IDL and its tiny community.
According to the 2014 TIOBE Index\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
 it is one of the top ten most popular programming languages in the world 
and is widely used outside of scientific fields in areas like web development, education 
and `big data' analytics.

The development of a package such as SunPy is made possible by the 
rich ecosystem of scientific packages available in Python. Core packages such 
as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language,
e.g., array manipulation, core numerical algorithms and visualisation. 
Building upon these foundations, packages such as \texttt{astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide more domain-specific functionality.

SunPy is designed to be a clean, simple-to-use, and well-structured 
package that provides the \textit{core} tools for solar data analysis. 
The primary focus of SunPy's early development are 
specialised data types that allow the acquisition, processing and 
visualisation of all types of solar data.

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the project's development model, community aspects of the 
SunPy project, as well as future plans. The latest release of SunPy, version 0.4,
can be downloaded from \href{http://sunpy.org}{www.sunpy.org} or can be
installed using the \href{http://pypi.python.org/pypi}{Python package index}.

>>>>

Science is driven by the analysis of data of growing variety and complexity.
Modern advances in sensor
technology, combined with the availability of inexpensive storage, has led to
rapid increases in the sheer amount of data available to scientists in almost
every discipline.  Solar physics is no exception to this trend. For example,
NASA's \textit{Solar Dynamics Observatory} (\textit{SDO}) spacecraft, launched in February 2010,
produces over 1 TB of catalogued and stored data per day \citep{aia}. Managing and
analysing this mountain of data
requires increasingly sophisticated software tools.
These tools should be robust, be easy of use and modify, have a
transparent development history, and conform
to standards.  Software with these qualities provide a strong
foundation and can support the needs of the community as data
volumes grow and science questions evolve.
%schriste - the opening paragraph still needs work
%ji - I had a go.
%schriste - looks good, made a small fix

The SunPy project aims to provide a free, open-source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and scientific Python packages. Python is a general-purpose,
powerful and easy-to-learn high-level programming language.
% RJH: Can we cut the next sentence?  It seems superfluous and top 10 is rather
% meaningless.
% schriste - i think it's important. it contrasts nicely with IDL and its tiny 
%community.
%SJM: I think it should stay, it makes a point.
According to the 2014 TIOBE Index\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
 it is one of the top ten most popular programming languages in the world 
and is widely used outside of scientific fields in areas like web development, education 
and `big data' analytics.

The development of a package such as SunPy is made possible by the 
rich ecosystem of scientific packages available in Python. Core packages such 
as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language:
e.g., array manipulation, core numerical algorithms and visualisation. 
Building upon these foundations, packages such as \texttt{astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide more domain-specific functionality.

SunPy is designed to be a clean, simple-to-use, and well-structured 
package that provides the \textit{core} tools for solar data analysis. 
The primary focus of SunPy's early development are 
specialised data types that enable the acquisition, processing and 
visualisation of all types of solar data.

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the project's development model, community aspects of the 
SunPy project, and future plans for the project. The latest release of SunPy, version 0.4,
can be downloaded from \href{http://sunpy.org}{www.sunpy.org} or can be
installed using the \href{http://pypi.python.org/pypi}{Python package index}.

### 2014-01-30 2-1-Map.tex

\subsection{Map}\label{ssec:map}
The Map data type stores 2D spatial data, such as images of the Sun and 
inner heliosphere. It provides a wrapper around a \texttt{numpy} data array and 
its associated spatial coordinates and other metadata. The \texttt{Map} class 
provides methods for typical operations on 2D data, such as rotation and 
re-sampling, as well as visualisation functionality.
The \texttt{Map} class also provides a convenient interface for loading data 
from a variety of sources, including a FITS file as shown in 
Listing~\ref{code:aia_1}.

The architecture of the \texttt{Map} class consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. 
\texttt{NDData} is a generic wrapper around an \texttt{ndarray} with a 
\texttt{meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does 
not yet make full use of its capabilities, but this inheritance structure 
provides for future integration of \texttt{astropy}. In order to provide 
instrument- or detector-specific integration, \texttt{GenericMap} is designed
to be subclassed. Each subclass of \texttt{GenericMap} can register 
with the \texttt{Map} creation factory, which will then automatically return an instance
of the specific \texttt{GenericMap} subclass dependant upon the data provided. 
SunPy v0.4 has \texttt{Map} specialisations for the following instruments: 
\textit{YOHKOH/SXT}, \textit{SOHO/EIT} and \textit{LASCO}, \textit{RHESSI}, 
\textit{STEREO/EUVI} and \textit{COR}, \textit{HINODE/XRT},
\textit{PROBA2/SWAP}, \textit{SDO/AIA} and \textit{HMI}, 
and \textit{IRIS} SJI frames. 
% RJH: Instruments should be consistent throughout (italics, caps, etc).  Also,
% usually citations are provided for instruments.

The \texttt{Map} class stores standard metadata retrieved from the header of 
the image file in the \texttt{meta} attribute and provides convenience 
properties for commonly accessed metadata: e.g., \texttt{instrument}, 
\texttt{wavelength} or \texttt{coordinate\_system}. 
Listing \ref{code:aia_1} demonstrates the quick-look functionality of 
\texttt{Map}; see Section \ref{subsec:Viz} for more discussion of plotting.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Example of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The map is created from an \textit{AIA} FITS file,
and a quick-view plot is created.}
\label{code:aia_1}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import matplotlib.pyplot as plt
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
...                         composite=True)
>>> compmap.set_levels(1, range(0,50,5), percent=True)
>>> compmap.set_colors(1, 'Reds_r')
#Plot the result and crop
>>> ax = plt.subplot()
>>> compmap.plot()
>>> ax.axis([200, 600, -600, -200])
>>> plt.show()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\end{center}
\caption{Example showing a \texttt{CompositeMap} plot, with RHESSI data composed with SDO/AIA data, and the integration with the \texttt{matplotlib.pyplot} interface.}
\label{code:compmap_1}
\end{listing}

%schriste - stuart, i agree if you want to remove the following two listings.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> compmap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_cube_controls}
\end{center}
\caption{Example showing creation of a \texttt{MapCube} from a glob file search. The 
resultant plot makes use of \texttt{matplotlib}'s interactive widgets to allow scrolling 
through the \texttt{MapCube}.}
\label{code:mapcube_1}
\end{listing}

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> from matplotlib import animation
>>> mapc = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> anim = mapc.plot()
>>> Writer = animation.writers['ffmpeg']
>>> writer = Writer(fps=10, metadata=dict(artist='SunPy'), bitrate=1800)
>>> anim.save('aia_cube.ogv')
\end{minted}
% In the above example, is Writer, writer clear? 
% Could we add more metadata to the movie? obs_date, instrument, etc...?
\caption{Example showing how to save a video animation from a \texttt{MapCube}, using 
\texttt{matplotlib}'s animation framework.}
\label{code:mapcube_2}
\end{listing}
% RJH: If this is kept, the 'artist' bit should be explained.

>>>>

\subsection{Map}\label{ssec:map}
The Map data type stores 2D spatial data, such as images of the Sun and 
inner heliosphere. It provides: a wrapper around a \texttt{numpy} data array; 
the images associated spatial coordinates; and other metadata. The \texttt{Map} 
class provides methods for typical operations on 2D data, such as rotation and 
re-sampling, as well as visualisation functionality.
The \texttt{Map} class also provides a convenient interface for loading data 
from a variety of sources, including a FITS file as shown in 
Listing~\ref{code:aia_1}.

The architecture of the \texttt{Map} class consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. 
\texttt{NDData} is a generic wrapper around a \texttt{numpy.ndarray} with a 
\texttt{meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does 
not yet make full use of its capabilities, but this inheritance structure 
provides for future integration with \texttt{astropy}. In order to provide 
instrument- or detector-specific integration, \texttt{GenericMap} is designed
to be subclassed. Each subclass of \texttt{GenericMap} can register 
with the \texttt{Map} creation factory, which will then automatically return an instance
of the specific \texttt{GenericMap} subclass dependant upon the data provided. 
SunPy v0.4 has \texttt{Map} specialisations for the following instruments: 
\textit{Yohkoh}/SXT, \textit{SOHO}/EIT and LASCO, \textit{RHESSI}, 
\textit{STEREO}/EUVI and COR, \textit{Hinode}/XRT,
\textit{PROBA2}/SWAP, \textit{SDO}/AIA and HMI, 
and \textit{IRIS} SJI frames. 
% RJH: Instruments should be consistent throughout (italics, caps, etc).  Also,
% usually citations are provided for instruments.
% dps: I would agree to add the citations if we were using the instrument in 
% the paper, but I don't think this is the case here.

The \texttt{Map} class stores standard metadata retrieved from the header of 
the image file in the \texttt{meta} attribute and provides convenience 
properties for commonly accessed metadata: e.g., \texttt{instrument}, 
\texttt{wavelength} or \texttt{coordinate\_system}. 
Listing \ref{code:aia_1} demonstrates the quick-look functionality of 
\texttt{Map}; see Section \ref{subsec:Viz} for additional discussion of plotting.
%dps: See section viz does not apply any more, right? The examples have been 
%removed now.
%SJM: the example is gone, but the section remains
&ays: it does remain, but it's so short it's not worth referring to

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import map
>>> aiamap = map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Example of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The map is created from an \textit{SDO}/AIA FITS file,
and a quick-view plot is created.}
\label{code:aia_1}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import map
>>> import matplotlib.pyplot as plt
>>> compmap = map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
...                   composite=True)
>>> compmap.set_levels(1, range(0,50,5), percent=True)
>>> compmap.set_colors(1, 'Reds_r')
#Plot the result and crop
>>> ax = plt.subplot()
>>> compmap.plot()
>>> ax.axis([200, 600, -600, -200])
>>> plt.show()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\end{center}
\caption{Example showing a \texttt{CompositeMap} plot, with RHESSI data composited
with \textit{SDO}/AIA data, and the integration with the \texttt{matplotlib.pyplot} interface.}
\label{code:compmap_1}
\end{listing}

%schriste - stuart, i agree if you want to remove the following two listings.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import map
>>> compmap = map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> compmap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_cube_controls}
\end{center}
\caption{Example showing creation of a \texttt{MapCube} from a glob file search. The 
resultant plot makes use of \texttt{matplotlib}'s interactive widgets to allow scrolling 
through the \texttt{MapCube}.}
\label{code:mapcube_1}
\end{listing}

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import map
>>> from matplotlib import animation
>>> mapc = map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> anim = mapc.plot()
>>> Writer = animation.writers['ffmpeg']
>>> writer = Writer(fps=10, metadata=dict(artist='SunPy'), bitrate=1800)
>>> anim.save('aia_cube.ogv')
\end{minted}
% In the above example, is Writer, writer clear? 
% Could we add more metadata to the movie? obs_date, instrument, etc...?
\caption{Example showing how to save a video animation from a \texttt{MapCube} using 
\texttt{matplotlib}'s animation framework.}
\label{code:mapcube_2}
\end{listing}
% RJH: If this is kept, the 'artist' bit should be explained.

### 2014-01-30 2-3-Spectra.tex

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work
%ji - there is no example of a spectrum
%dps - I need to think in one...
%schriste - can't we just take a slice off of the spectram to make a spectrum?

SunPy aims to provide broad support for solar spectroscopy
instruments.  The variety and complexity of these instruments and
their resultant datasets makes this a challenging goal.  The \texttt{spectra} module implements a
\texttt{Spectrum} class for 1D data (intensity as a function of frequency) and a
\texttt{Spectrogram} class for 2D data (intensity as a function of time and
frequency).  Each of these classes use a \texttt{numpy.ndarray} class
as its \texttt{data} attribute.  These two classes were implemented
by funding provided by the Astrophysics Research Group at Trinity
College Dublin, Ireland.

As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities. The common functionality provided by the base \texttt{Spectrogram} class includes
joining different time ranges and frequencies, performing frequency-dependent background subtraction,
and convenient visualization and sampling of the data.
Currently, the \texttt{Spectrogram} class supports radio spectrograms from the 
\href{http://www.e-callisto.org/}{e-Callisto}
solar radio spectrometer network and STEREO/SWAVES spectrograms.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{from\_range()} function, the object merges all the downloaded
files into a single spectrogram, across time and frequency.
In the example shown, data is provided in two frequency ranges, 
20--90\,MHz and 55--355\,MHz.  Since the data is not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis for a better analysis.  The example also demonstrates
the implemented background subtraction method.
% RJH: Which subtraction method?  If the specifics are irrelevant, cut it.
%ji to DPS - what is the linearisation that is happening here?

@@

% Download Callisto
% Merge multiple time-ranges / frequencies (just work from downlad!)
% Merge callisto with swaves

>>>>

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work
%ji - there is no example of a spectrum
%dps - I need to think in one...
%schriste - can't we just take a slice off of the spectram to make a spectrum?
%dps - Spectrum does not really do much, just peek and plot. I don't think
% it brings anything to the content.
SunPy aims to provide broad support for solar spectroscopy
instruments.  The variety and complexity of these instruments and
their resultant datasets makes this a challenging goal.  The \texttt{spectra} module implements a
\texttt{Spectrum} class for 1D data (intensity as a function of frequency) and a
\texttt{Spectrogram} class for 2D data (intensity as a function of time and
frequency).  Each of these classes use a \texttt{numpy.ndarray} object
as its \texttt{data} attribute.  These two classes were implemented
by funding provided by the Astrophysics Research Group at Trinity
College Dublin, Ireland.

As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities. The common functionality provided by the base \texttt{Spectrogram} class includes
joining different time ranges and frequencies, performing frequency-dependent background subtraction,
and convenient visualization and sampling of the data.
Currently, the \texttt{Spectrogram} class supports radio spectrograms from the 
\href{http://www.e-callisto.org/}{e-Callisto}
solar radio spectrometer network and \textit{STEREO}/SWAVES spectrograms.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{from\_range()} function, the object merges all the downloaded
files into a single spectrogram, across time and frequency.
In the example shown, data is provided in two frequency ranges:
20--90\,MHz and 55--355\,MHz.  Since the data is not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis to assist analysis.  The example also demonstrates
the implemented background subtraction method, which calculates
a constant background over time for each frequency channel.
% RJH: Which subtraction method?  If the specifics are irrelevant, cut it.
% RJH: is this better?
%ji to DPS - what is the linearisation that is happening here?
%ji- the linearisation consist in the frequency axis being linear and the 
% data being streched with it.  The frequency normally are not evenly spaced
%ays: I deleted the phrase "based on the values closer to the average" because I don't understand what that means

@@

### 2014-01-30 2-Data_Types.tex
comments and some little fixings

These classes allow access to the data
along with associated metadata and provide appropriate convenience functions to
enable analysis and visualisation. For each of these classes, the data is
stored in the \texttt{data} attribute while the metadata is stored 
in the \texttt{meta} attribute. (Curently, \texttt{LightCurve} does 
not support the \texttt{meta} attribute.) 
% schriste - can't we just quickly fix this? should be easy enough.
It is possible to instantiate the
data types from various
different sources: e.g., files, URLs, arrays.  
%schriste - i removed time ranges as that just looks up a url.
In order to provide instrument-specific specialisation, all the core SunPy classes 
support subclassing; e.g., \texttt{Map} has an \texttt{AIAMap} 
sub-type for data from the SDO/AIA instrument. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick 
plotting, and \texttt{plot()}, for plotting with more fine-grained control that 
integrates with \texttt{matplotlib}.

@@

\subsection{Visualisation}
\label{subsec:Viz}
As demonstrated in this section, the core SunPy datatypes 
include visualisation methods that are tailored to that data type. 
These visualisation methods all currently utilise the \texttt{matplotlib} 
package, and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

>>>>

These classes allow access to the data
and associated metadata and provide appropriate convenience functions to
enable analysis and visualisation. For each of these classes, the data is
stored in the \texttt{data} attribute while the metadata is stored 
in the \texttt{meta} attribute. (Curently, \texttt{LightCurve} does 
not support the \texttt{meta} attribute.) 
% schriste - can't we just quickly fix this? should be easy enough.
% dps - Spectrogram does not contain meta either! I would suggest
%(Curently, this is just supported by Map, but the other data types will 
%include such attribute soon.)
%SJM: oh what, drop it all together.
It is possible to instantiate the
data types from various
different sources: e.g., files, URLs, arrays.  
%schriste - i removed time ranges as that just looks up a url.
In order to provide instrument-specific specialisation, the core SunPy classes 
make use of subclassing; e.g., \texttt{Map} has an \texttt{AIAMap} 
sub-type for data from the \textit{SDO}/AIA instrument. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick 
plotting, and \texttt{plot()}, for plotting with more fine-grained control that 
integrates with \texttt{matplotlib}.

@@

\subsection{Visualisation}
\label{subsec:Viz}
As demonstrated in this section, the core SunPy data types 
include visualisation methods that are tailored to each data type. 
These visualisation methods all currently utilise the \texttt{matplotlib} 
package, and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

### 2014-01-30 3-4-Helioviewer.tex

SunPy provides the ability to download images hosted by the
\href{http://helioviewer.org}{Helioviewer Project}.  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces.  The primary interface to Helioviewer is a
google-map-like website which allows users to browse images of the Sun
taken by a variety of instruments. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data as JPEG2000 format (for further
details on the Helioviewer Project see \citep{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> hv.download_png('2099/01/01', 6,
...                 "[SDO,AIA,AIA,304,1,100],[SDO,AIA,AIA,193,1,50]"+
...                 "[SOHO,LASCO,C2,white-light,1,100]",
...                 x0=0, y0=0, width=768, height=768)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_overlay_example}
\end{center}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

The first argument is the requested time of the image.  Helioviewer
selects images closest to the requested time.  In this case, the
requested time is in the future and so Helioviewer will find the most
recent available images from each source.  The second argument refers
to the image resolution in arcseconds per pixel (larger values mean
lower resolution).  The third argument is a comma-delimited string of
the three requested image layers, the details of which are enclosed
in parentheses. The image layers are described using the observatory
$\rightarrow$ instrument $\rightarrow$ detector $\rightarrow$
measurement combination described above, along with the final two
numbers that denote the visibility and the opacity of the image layer
respectively (1/0 is visible/invisible with opacity in the range
$0\rightarrow100$, with $100$ meaning fully opaque).  The quantities
\texttt{x0} and \texttt{y0} are the $x$ and $y$ centre points about
which to centre the image (measured in helio-projective cartesian
co-ordinates) and the \texttt{width} and \texttt{height} are the pixel
values for the image dimensions.

@@

As noted above, Helioviewer JPEG2000 files contain metadata that allow
positioning of the image data.  There is sufficient metadata in each
file to permit the creation of a SunPy map object (see Section
\ref{ssec:map}) from a Helioviewer JPEG2000 file.  This allows image
data to be manipulated in the same way as any other map object.

@@

Listing \ref{code:downloadjp2} demonstrates the querying, downloading,
reading and conversion of a Helioviewer JPEG2000 file into a SunPy map
object.  This functionality allows users to visualise and manipulate
Helioviewer-supplied image data in an identical fashion to a SunPy map
object generated from FITS data (see Section \ref{ssec:map}).

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> filepath = hv.download_jp2('2012/07/05 00:30:00',
...                            observatory='SDO',
...                            instrument='HMI', detector='HMI', 
...                            measurement='continuum')
>>> from sunpy.map import Map
>>> Map(filepath).submap([200,550],[-400,-200]).peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy map object.}
\label{code:downloadjp2}
\end{listing}

>>>>

SunPy provides the ability to download images hosted by the
\href{http://helioviewer.org}{Helioviewer Project}.  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces.  The primary interface to Helioviewer is a
Google-Maps-like website which allows users to browse images of the Sun
taken by a variety of instruments. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data in JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and thus can be used to quickly
visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 6,
...                 "[SDO,AIA,AIA,304,1,100],[SDO,AIA,AIA,193,1,50],"+
...                 "[SOHO,LASCO,C2,white-light,1,100]",
...                 x0=0, y0=0, width=768, height=768)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_overlay_example}
\end{center}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

The first argument is the requested time of the image.  Helioviewer
selects images closest to the requested time.  In this case, the
requested time is in the future and so Helioviewer will find the most
recent available images from each source.  The second argument refers
to the image resolution in arcseconds per pixel (larger values mean
lower resolution).  The third argument is a comma-delimited string of
the three requested image layers, the details of which are enclosed
in parentheses. The image layers are described using the observatory
$\rightarrow$ instrument $\rightarrow$ detector $\rightarrow$
measurement combination described above, along with the final two
numbers that denote the visibility and the opacity of the image layer
respectively (1/0 is visible/invisible with opacity in the range
$0\rightarrow100$, with $100$ meaning fully opaque).  The quantities
\texttt{x0} and \texttt{y0} are the $x$ and $y$ centre points about
which to centre the image (measured in helio-projective cartesian
coordinates), and the \texttt{width} and \texttt{height} are the pixel
values for the image dimensions.

@@

As noted above, Helioviewer JPEG2000 files contain metadata that allow
positioning of the image data.  There is sufficient metadata in each
file to permit the creation of a SunPy \texttt{Map} object (see Section
\ref{ssec:map}) from a Helioviewer JPEG2000 file.  This allows image
data to be manipulated in the same way as any other map object.

@@

Listing \ref{code:downloadjp2} demonstrates the querying, downloading,
reading and conversion of a Helioviewer JPEG2000 file into a SunPy map
object.  This functionality allows users to visualise and manipulate
Helioviewer-supplied image data in an identical fashion to a SunPy \texttt{Map}
object generated from FITS data (see Section \ref{ssec:map}).

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> filepath = hv.download_jp2('2012/07/05 00:30:00',
...                            observatory='SDO',
...                            instrument='HMI', detector='HMI', 
...                            measurement='continuum')
>>> from sunpy import map
>>> map.Map(filepath).submap([200,550],[-400,-200]).peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy \texttt{Map} object.}
\label{code:downloadjp2}
\end{listing}

### 2014-01-30 3-5-Database.tex

The database sub-package makes use of a database supported by
\href{http://www.sqlalchemy.org}{SQLAlchemy}. This library was chosen
since it supports many SQL dialects. 
If SQLite is selected, the database is stored as a single file which is
created automatically. A server-based database, on the other hand, could be used
by a collaborators who work together on the
same data from different computers: a central database server stores all data and the clients connect to
it to read or write data.

The database can store and manage all data that can be read via SunPy's 
\texttt{io} subpackage, and direct integration with the \textsc{VSO} 
subpackage is supported.
It is also possible to manually add file or directory entries. The package also provides
a unified data search via the \texttt{fetch()} method, which includes both local files
and files on the \textsc{VSO}. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database, the file is scanned and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom meta-data to 
each database entry such as keywords, comments, and favouring, as well as 
%dps: favouring is not what we want to say here...
querying the full metadata (i.e., FITS header) of each entry.

@@

A very special feature of the database package is the support of \texttt{undo}
and \texttt{redo} operations. This is particularly convenient in
interactive sessions to easily revert accidental operations. 
This feature will also be desirable when creating a GUI frontend for this package.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import vso
>>> from sunpy.database import Database
>>> database = Database('sqlite:///')
>>> database.download(
...     vso.attrs.Time('2012-08-05', '2012-08-05 00:00:05'),
...     vso.attrs.Instrument('AIA'))
>>> len(database)
2
>>> from sunpy.database.tables import display_entries
>>> print display_entries(
...     database,
...     ['id', 'observation_time_start', 'wavemin', 'wavemax'])
id observation_time_start wavemin wavemax
-- ---------------------- ------- -------
1  2012-08-05 00:00:01    9.4     9.4    
2  2012-08-05 00:00:02    33.5    33.5   
\end{minted}
\caption{Example usage of the database sub-package.}
\label{code:database}
\end{listing}

>>>>

The \texttt{database} subpackage makes use of a database supported by
\href{http://www.sqlalchemy.org}{\texttt{SQLAlchemy}}. This library was chosen
since it supports many SQL dialects. 
If SQLite is selected, the database is stored as a single file which is
created automatically. A server-based database, on the other hand, could be used
by a collaborators who work together on the
same data from different computers: a central database server stores all data and the clients connect to
it to read or write data.

The database can store and manage all data that can be read via SunPy's 
\texttt{io} subpackage, and direct integration with the \texttt{vso} 
module is supported.
It is also possible to manually add file or directory entries. The package also provides
a unified data search via the \texttt{fetch()} method, which includes both local files
and files on the \textsc{VSO}. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database, the file is scanned for metadata,
and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom metadata to 
each database entry such as keywords, comments, and favouring, as well as 
%dps: favouring is not what we want to say here...
%JI: ``assigning favourites'' - is this what is meant?
%ays: i'll let someone else fix this
querying the full metadata (e.g., FITS header) of each entry.

@@

A useful feature of the database package is the support of \texttt{undo}
and \texttt{redo} operations. This is particularly convenient in
interactive sessions to easily revert accidental operations. 
This feature will also be desirable when creating a GUI frontend for this package.
%JI - revert accidental operations of any kind?

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import vso
>>> from sunpy.database import Database
>>> database = Database('sqlite:///')
>>> database.download(
...     vso.attrs.Time('2012-08-05', '2012-08-05 00:00:05'),
...     vso.attrs.Instrument('AIA'))
>>> len(database)
2
>>> from sunpy.database.tables import display_entries
>>> print display_entries(
...     database,
...     ['id', 'observation_time_start', 'wavemin', 'wavemax'])
id observation_time_start wavemin wavemax
-- ---------------------- ------- -------
1  2012-08-05 00:00:01    9.4     9.4    
2  2012-08-05 00:00:02    33.5    33.5   
\end{minted}
\caption{Example usage of the \texttt{database} subpackage.}
\label{code:database}
\end{listing}

### 2014-01-30 3-Retrieval.tex

Several well-developed resources currently exist within heliophysics which provide event databases, remote access and data retrieval for a large number of data sources. As a community-based project, SunPy provides support for these resources via the \texttt{sunpy.net} package. In the following subsections, we describe a number of these resources and introduce current methods for accessing them within SunPy.

>>>>

Several well-developed resources currently exist within heliophysics which provide event databases, 
remote access and data retrieval for a large number of data sources. As a community-based project, 
SunPy provides support for these resources via the \texttt{net} subpackage. 
In the following subsections, we describe a number of these resources and 
how to access them using SunPy.
